---
title: "Introduction to MDFA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to MDFA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MDFA)
```

## Simulating data with a known factor structure

```{r}
set.seed(1)

p <- 20
m <- 3
n <- 100

f_size <- p / m
Lam_ast <- matrix(0, nrow = p, ncol = m)
Lam_ast[0 * f_size + (1:f_size), 1] <- 0.9
Lam_ast[1 * f_size + (1:f_size), 2] <- 0.8
Lam_ast[2 * f_size + (1:f_size), 3] <- 0.6
class(Lam_ast) <- "loadings"

uniq_var <- 1 - rowSums(Lam_ast^2)
Sig_ast <- Lam_ast %*% t(Lam_ast) + diag(uniq_var)

data <- mvtnorm::rmvnorm(n = n, mean = rep(0, p), sigma = Sig_ast)
```

## Fitting MDFA (no bootstrap)

```{r}
res_als <- MDFA(
  data = data, nfactors = m, SMC = TRUE, alg = "ALS",
  n.iter = 1, tol = 1e-5, trace = 0, REPORT = 0, rotate = "varimax"
)
res_als$loadings

res_bfgs <- MDFA(
  data = data, nfactors = m, SMC = TRUE, alg = "BFGS",
  n.iter = 1, tol = 1e-5, trace = 0, REPORT = 0, rotate = "varimax"
)
res_bfgs$loadings
```

## Fitting MDFA with bootstrap confidence intervals

`n.iter` controls the number of non-parametric bootstrap replications used
to compute standard errors and confidence intervals for the loadings. Larger
values of `n.iter` (e.g. `10^3`, as in the original script this package is
based on) give more stable intervals but take proportionally longer; this
vignette uses a small value so it builds quickly.

```{r}
res_ci <- MDFA(
  data = data, nfactors = m, SMC = TRUE, alg = "ALS",
  n.iter = 30, tol = 1e-5, trace = 0, REPORT = 0, rotate = "varimax",
  num_cores = 1
)

res_ci$cis$means
res_ci$cis$sds
```

## Fitting directly from a covariance matrix

Internally, `MDFA()` computes `cov(data)` and passes it to a non-exported
helper, `MDFA_cov()`, which does the actual point estimation without any
bootstrapping. If you already have a covariance (or correlation) matrix and
don't need bootstrap standard errors, you can call it directly with `:::`:

```{r}
S <- cov(data)
res_direct <- MDFA_cov(covar = S, nfactors = m, alg = "ALS", trace = 0, REPORT = 0, rotate = "varimax")
res_direct$loadings
```
